---
title: 'Biodiversity Sampling: Exposed vs Sheltered'
author: "Andrew Bickell, Gabrielle Languedoc, Joan Moreaux and Mara Bohm"
date: "30/10/2021"
output:
  pdf_document:
    includes:
      in_header: file.tex
---

## Hypothesis

We hypothesize that there will be greater biodiversity found in the sheltered site because it experiences less hydrodynamic stress.

## Data Collection

We collected data from both wave-shelteredand an wave-exposed sites, at Scott's Bay, Bamfield. We used 10 by 10 quadrats and selected 10 random squares, in which we counted species richness and abundance. 

## Analysis

```{r, echo=FALSE, message=FALSE}
library(vegan)
library(tidyverse)
```

```{r, echo=FALSE, message=FALSE}
data <- read.csv("data-entry-final.csv")

# Manipulating the data

data_analysis <- data %>% 
  select(-c(abiotic_hab, biotic_hab)) %>% 
  mutate(quadrat = as.factor(quadrat)) %>% 
  group_by(vertical_transect, quadrat, species)%>%
  mutate(n_per_quadrat = sum(n)*10) %>%
  mutate(site = ifelse(vertical_transect %in% c(1,2,3), "sheltered", "exposed")) %>% 
  arrange(vertical_transect, quadrat) %>%
  select(-c(n,square)) %>% 
  na.omit %>%
  distinct() %>%
  group_by(vertical_transect, quadrat) %>%
  mutate(quadrat_prop = n_per_quadrat / sum(n_per_quadrat))

quad_richness <- data %>%
  drop_na(n) %>%
  group_by(vertical_transect,quadrat) %>%
  summarize(richness = length(unique(species))) %>%
  mutate(site = ifelse(vertical_transect %in% c(1,2,3), "sheltered", "exposed"))
```

```{r}
richness_plot <- ggplot(data = quad_richness) +
  geom_boxplot(aes(x = site, y = richness, fill = site)) +
  labs(x = "Site", y = "Species Richness")+
  theme_grey()+
  theme(legend.position = 'none')
    
richness_plot
```

## Results and Conclusion
