---
title: 'Biodiversity Sampling: Exposed vs Sheltered'
author: "Andrew Bickell, Gabrielle Languedoc, Joan Moreaux and Mara Bohm"
date: "01/11/2021"
output:
  pdf_document:
    includes:
      in_header: file.tex
---

## Hypothesis

We hypothesize that there will be greater biodiversity found in the sheltered site because it experiences less hydrodynamic stress.

## Data Collection

We collected data from both wave-sheltered and an wave-exposed sites, at Scott's Bay, Bamfield. We counted species richness and abundance within ten haphazardly selected 10 cm by 10 cm quadrats along two parallel transect lines per site. 

## Analysis

We compared Shannon-Weiner diveristy indexs (H'), species richness and, species evenness between the exposed and sheltered sites. We ran the statistical analysis ANOVA to determine the significance of our results. 


```{r, echo=FALSE, message=FALSE}
library(vegan)
library(here)
library(scales)
library(viridis)
library(tidyverse)
```

```{r, echo=FALSE, message=FALSE}
data <- read.csv("data-entry-final.csv")

# Manipulating the data

data_analysis <- data %>% 
  select(-c(abiotic_hab, biotic_hab)) %>% 
  mutate(quadrat = as.factor(quadrat)) %>% 
  group_by(vertical_transect, quadrat, species)%>%
  mutate(n_per_quadrat = sum(n)*10) %>%
  mutate(site = ifelse(vertical_transect %in% c(1,2,3), "sheltered", "exposed")) %>% 
  select(-c(n,square)) %>% 
  na.omit %>%
  distinct() %>%
  group_by(vertical_transect, quadrat) %>%
  mutate(pi_quadrat = n_per_quadrat / sum(n_per_quadrat)) %>% 
  mutate(pi_log = log(pi_quadrat)) %>% 
  mutate(pi_times_pi_log = -(pi_quadrat * pi_log)) %>% 
  select(-c(pi_quadrat, pi_log))

# Diversity calculation

shannon_diversity <- data_analysis %>% 
  group_by(vertical_transect, quadrat) %>%
  summarize(sum(pi_times_pi_log, na.rm = FALSE)) %>%
  rename("shannon_div" = "sum(pi_times_pi_log, na.rm = FALSE)") %>%
  mutate(site = ifelse(vertical_transect %in% c(1,2,3), "sheltered", "exposed"))

# Richness calculation

quad_richness <- data %>%
  drop_na(n) %>%
  group_by(vertical_transect,quadrat) %>%
  summarize(richness = length(unique(species))) %>%
  mutate(site = ifelse(vertical_transect %in% c(1,2,3), "sheltered", "exposed"))

# Evenness calculation

shannon_evenness <- shannon_diversity %>% 
  cbind(richness = quad_richness$richness) %>% 
  mutate(H_max = log(richness)) %>%
  mutate(evenness = shannon_div/H_max)
```

## Results and Conclusions

<<<<<<< HEAD
```{r, fig.align = 'left', out.width = "80%", out.height = "80%", fig.cap = "Species richness present at the exposed and sheltered sites at Scott's Bay, Bamfield, Canada. Determined using 10 quadrats haphazardly sampled along two parallel transect lines per site", fig.cap.align = 'left', echo=FALSE, warning=FALSE, message=FALSE}
=======
Our analysis found that there was no significant difference between the Shannon-Weiner indexes of the exposed and sheltered sites. The p-value for this anova test was 0.218. Similarly, the p value of the richness and evenness anova analysis were 0.43 and 0.78 respectively. In all cases, we cannot reject the null hypothesis that there is no relationship between site location and diversity, richness and evenness (Table 1). 

```{r, fig.align = 'left', out.width = "80%", out.height = "80%", fig.cap = "Hello.", fig.cap.align = 'left', echo=FALSE, warning=FALSE, message=FALSE}
>>>>>>> main
richness_plot <- ggplot(data = quad_richness) +
  geom_boxplot(aes(x = site, y = richness, fill = site)) +
  labs(x = "Site", y = "Species Richness")+
  theme_test()+
  theme(legend.position = 'none') +
  scale_fill_manual(values = c('#A8327DFF', '#FA7F5EFF'))


richness_plot # COLE: how to export?
ggsave(here("./images/richness_plot.png"), richness_plot)

```
```{r, echo=FALSE, results=FALSE}
## Statistical Analysis
anova_test_rich <- aov(richness~ site, data = quad_richness)
summary(anova_test_rich)
```

```{r, fig.align = 'left', out.width = "80%", out.height = "80%", fig.cap = "fig.cap = "H' value of Shannon-Wiener Diversity Index present at the exposed and sheltered sites at Scott's Bay, Bamfield, Canada. Determined using 10 quadrats haphazardly sampled along two parallel transect lines per site.", fig.cap.align = 'left', echo=FALSE, warning=FALSE, message=FALSE}
shannon_diversity_plot <- ggplot(data = shannon_diversity) +
  geom_boxplot(aes(x = site, y = shannon_div, fill = site)) +
  labs(x = "Site", y = "Diversity Index")+
  theme_test()+
  theme(legend.position = 'none')+
  scale_fill_manual(values = c('#A8327DFF', '#FA7F5EFF'))

shannon_diversity_plot
ggsave(here("./images/diversity_plot.png"), shannon_diversity_plot)
```

\newpage 

```{r, echo=FALSE, results=FALSE}
## Statistical Analysis
anova_test_div <- aov(shannon_div ~ site, data = shannon_diversity)
summary(anova_test_div)
```

```{r, fig.align = 'left', out.width = "80%", out.height = "80%", fig.cap = "Diversity evenness values of the exposed and sheltered sites at Scott's Bay, Bamfield, Canada. Determined using 10 quadrats haphazardly sampled along two parallel transect lines per site.", fig.cap.align = 'left', echo=FALSE, warning=FALSE, message=FALSE}
shannon_evenness_plot <- ggplot(data = shannon_evenness) +
  geom_boxplot(aes(x = site, y = evenness, fill = site)) +
  labs(x = "Site", y = "Eveness Index")+
  theme_test()+
  theme(legend.position = 'none')+
  scale_fill_manual(values = c('#A8327DFF', '#FA7F5EFF'))

shannon_evenness_plot
ggsave(here("./images/evenness_plot.png"), shannon_evenness_plot)
```

```{r, echo=FALSE, results=FALSE}
## Statistical Analysis
anova_test_even <- aov(evenness ~ site, data = shannon_evenness)
summary(anova_test_even)
```

Table 1. 
| Index | P Value |
| ------- | --------- |
| Shannon-Weiner Diversity (H') | 0.218 |
| Species Richness | 0.43 |
| Species Evenness | 0.78 |
